---
title: "Analyses of the Perceptions of the Behavior that Led to Anger"
author: "Pooya Razavi"
date: "last knitted: `r Sys.time()`"
output: 
  html_document:
    theme: cosmo
    highlight: textmate
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---

This script represents the test of the hypotheses about the behavior that caused the participant to experience anger, and how the different aspects of the behavior vary between justified and unjustified anger prototypes.

```{r setup, include=FALSE, warning=FALSE}
#load libraries
package_list <- c("dplyr", "tidyr", "ggplot2", "MetBrewer")
lapply(package_list, require, character.only = TRUE)

#read in the data
df <- readxl::read_xlsx("C:/Users/pooya/Dropbox (University of Oregon)/Anger Dissertation/Prototype study analysis/ProcessedData_F21_W22_S22_F22.xlsx")

#Function to report correlation
cor_report <- function(cor_output) {
          r <- cor_output[["estimate"]] %>% round(2)
          df <- cor_output[["parameter"]] %>% round(1)
          ci_lb <- (cor_output[["conf.int"]])[1] %>% round(2)
          ci_ub <- (cor_output[["conf.int"]])[2] %>% round(2)
          original_p <- cor_output[["p.value"]] %>% round(3)
          p <- if_else(original_p >= .001, paste0("= ", as.character(original_p)), "< .001")
          
          print(paste0("r(", df, ") = ", r, " [", ci_lb, ", ", ci_ub, "], p ", p))
    }

#Function to report independent-samples t-test
  ind_ttest_report <- function(iv, dv) {
      ttest <- t.test(dv ~ iv)
      effect_size <- effectsize::cohens_d(dv ~ iv, pooled_sd = FALSE)
      t <- ttest[["statistic"]] %>% round(2)
      df <- ttest[["parameter"]] %>% round(1)
      original_p <- ttest[["p.value"]] %>% round(3)
      p <- if_else(original_p >= .001, paste0("= ", as.character(original_p)), "< .001")
      d <- effect_size[1,1] %>% round(2)    
      
      print(paste0("t(", df, ") = ", t, ", p ", p, ", d = ", d))
      
  }
  
#Function to report paired-samples t-test
    paired_ttest_report <- function(t1, t2) {
      ttest <- t.test(Pair(t1, t2) ~ 1)
      effect_size <- effectsize::cohens_d(Pair(t1, t2) ~ 1, pooled_sd = FALSE)
      t <- ttest[["statistic"]] %>% round(2)
      df <- ttest[["parameter"]] %>% round(1)
      original_p <- ttest[["p.value"]] %>% round(3)
      p <- if_else(original_p >= .001, paste0("= ", as.character(original_p)), "< .001")
      d <- effect_size[1,1] %>% round(2)    
      
      print(paste0("t(", df, ") = ", t, ", p ", p, ", d = ", d))
      
  }
  
  
knitr::opts_chunk$set(echo = TRUE)
```


```{r, data-exclusion}
#assigning values to factor levels
  df$NarrativeWritten <- as.factor(df$NarrativeWritten)
  df$NarrativeRelevant <- as.factor(df$NarrativeRelevant)
  df$Condition <- as.factor(df$Condition)
  
  levels(df$NarrativeWritten) <- c("No", "Yes")
  levels(df$NarrativeRelevant) <- c("No", "Yes", NA, NA) 
  levels(df$Condition) <- c("justified", "nonjustified", NA)

#drop cases following preregistration
  df1 <- df %>% 
    filter(NarrativeWritten != "No") %>% 
    filter(NarrativeRelevant != "No") %>% 
    filter(!is.na(Condition))

```


# Harm and threat

**Hypothesis 1:** Justified anger events (compared to the unjustified ones) involve a stronger perception of harm and threat to the self and to others.

**Hypothesis 2:** The difference between justified and unjustified anger in terms of perceptions of harm or threat is moderated by the target of harm, such that this difference is larger for harm to others (vs. harm to self). 

## Analyses 

```{r, harm-threat}
#First, examine the correlation between the 2 items for threat/harm to self and the 2 items for threat/harm to other. If r > .5, create composite scores by averaging the 2 items for each variable. 

  cor.test(df1$harm_you, df1$threat_you) %>% cor_report()
  cor.test(df1$harm_others, df1$threat_others) %>% cor_report()
  
#Since the items are strongly correlated (r > .5), we create 2 composites:

  #create the dataframe
  harm_threat_df <- df1 %>% 
                        select(ResponseId, Condition, harm_you, harm_others, threat_you, threat_others) %>% 
                        mutate(harm.threat_self = ((harm_you + threat_you) / 2),
                               harm.threat_other = ((harm_others + threat_others) / 2))
  
  #create a long dataframe
  harm_threat_df_long <- harm_threat_df %>% 
                              tidyr::pivot_longer(cols = c(harm.threat_self, harm.threat_other),
                                                  names_to = "target",
                                                  names_prefix = "harm.threat_",
                                                  values_to = "harm")

#The model: 
  #the main effect of type of anger (i.e., condition) corresponds to Hypothesis 1
  #the interaction effect corresponds to Hypothesis 2
  
  harm_threat_df_long %>% 
    rstatix::anova_test(data = .,
                        dv = harm,
                        wid = ResponseId,
                        between = Condition,
                        within = target) %>% 
    rstatix::get_anova_table()


#The difference between the two conditions in terms of harm/threat to self:
  t.test(harm.threat_self ~ Condition, data = harm_threat_df)
  effectsize::cohens_d(harm.threat_self ~ Condition, data = harm_threat_df)
  #For reporting: 
  ind_ttest_report(harm_threat_df$Condition, harm_threat_df$harm.threat_self)

#The difference between the two conditions in terms of harm/threat to others:
  t.test(harm.threat_other ~ Condition, data = harm_threat_df)
  effectsize::cohens_d(harm.threat_other ~ Condition, data = harm_threat_df)
  #For reporting: 
  ind_ttest_report(harm_threat_df$Condition, harm_threat_df$harm.threat_other)

```

## Figure

```{r, harm-threat-fig, fig.align='center', fig.width=5}

#getting mean and se for the graph
      descripts <- psych::describeBy(harm_threat_df_long$harm, 
                         group = list(harm_threat_df_long$Condition, harm_threat_df_long$target), mat = TRUE)   %>% 
                        select(group1, group2, mean, se) %>% 
                        transmute(Anger = as.factor(group1),
                                  Target = as.factor(group2),
                                  mean, se)
      levels(descripts$Anger) <- c("Justified", "Unjustified")
      levels(descripts$Target) <- c("Other", "Self")
      
#interaction plot - dot-whiskers and line plot
  harm_threat_p <- ggplot(
                descripts,
                aes(x = Anger, color = Target, group = Target, y = mean,
                    ymin = mean-se, ymax = mean+se)) + 
                geom_line(position = position_dodge(0.1)) +
                geom_errorbar(position = position_dodge(0.1), width = 0.12) +
                geom_point(aes(color = Target), position = position_dodge(0.1), size = .8) +
                ylim(1, 5) +
                ggtitle("Harm and Threat") +
                theme_bw() + 
                labs(y="Perception of Harm and Threat", x = "Anger") +
                theme(plot.title = element_text(hjust = 0.5), text = element_text(size = 10)) + 
                      scale_color_manual(values=met.brewer("Nizami", 2, direction=-1))      

    harm_threat_p
      #ggsave(plot = harm_threat_p, width = 5, height = 5, dpi = 300, filename = "harm_threat_p.png")
```

# Fairness and Justification

**Hypothesis:** The target's behavior is perceived as less fair and less justifiable in the justified (vs. unjustified) anger condition.

```{r, fair-justified}
#First, examine the correlation between the 2 items. If r > .5, proceed with creating a composite score. If not, conduct the analyses for each item separately.

  cor.test(df1$beh_fair, df1$beh_justified) %>% cor_report()

#Since items are strongly correlated (r > .5): 
  df1 <- df1 %>% 
            mutate(fair_just = ((beh_fair + beh_justified) / 2))
#The t-test
  df1 %>% 
      t.test(fair_just ~ Condition, data = .)
  
  df1 %>% 
      effectsize::cohens_d(fair_just ~ Condition, data = .)
  #For reporting
  ind_ttest_report(df1$Condition, df1$fair_just)
```

# Norm Violations

**Hypothesis 1:** The target’s behavior during justified anger events (vs. unjustified ones) is judged as a stronger norm violation. 

**Hypothesis 2:** The difference between justified and unjustified anger events is moderated by the type of norm violation, such that this difference is larger for injunctive norms (compared to descriptive norms). 

## Analyses

```{r, norms}
#create the dataframe
    norms_df <- df1 %>% 
                      select(ResponseId, Condition, how_acceptable, how_common) %>% 
                      mutate(injunctive = 6 - how_acceptable,
                             descriptive = 6 - how_common)
                      
                      

#create a long dataframe
    norms_df_long <- norms_df %>% 
                            tidyr::pivot_longer(cols = c(injunctive, descriptive),
                                                names_to = "norm_type",
                                                values_to = "violation")

#The model: 
    #The main effect of type of anger (i.e., condition) corresponds to Hypothesis 1
    #The interaction effect corresponds to Hypothesis 2
    
    norms_df_long %>% 
      rstatix::anova_test(data = .,
                          dv = violation,
                          wid = ResponseId,
                          between = Condition,
                          within = norm_type) %>% 
      rstatix::get_anova_table()


#The difference between the two conditions in terms of injunctive norms
    t.test(injunctive ~ Condition, data = norms_df)
    effectsize::cohens_d(injunctive ~ Condition, data = norms_df)
  #For reporting
    ind_ttest_report(norms_df$Condition, norms_df$injunctive)
    
#The difference between the two conditions in terms of descriptive norms
    t.test(descriptive ~ Condition, data = norms_df)
    effectsize::cohens_d(descriptive ~ Condition, data = norms_df)
  #For reporting
    ind_ttest_report(norms_df$Condition, norms_df$descriptive)
```

## Figure

```{r, norms-fig, fig.align='center', fig.width=5}

#getting mean and se for the graph
      descripts <- psych::describeBy(norms_df_long$violation, 
                         group = list(norms_df_long$Condition, norms_df_long$norm_type), mat = TRUE)   %>% 
                        select(group1, group2, mean, se) %>% 
                        transmute(Anger = as.factor(group1),
                                  Norms = as.factor(group2),
                                  mean, se)
      levels(descripts$Anger) <- c("Justified", "Unjustified")
      levels(descripts$Norms) <- c("Descriptive", "Injunctive")
      
#interaction plot - dot-whiskers and line plot
  norms_p <- ggplot(
                descripts,
                aes(x = Anger, color = Norms, group = Norms, y = mean,
                    ymin = mean-se, ymax = mean+se)) + 
                geom_line(position = position_dodge(0.1)) +
                geom_errorbar(position = position_dodge(0.1), width = 0.12) +
                geom_point(aes(color = Norms), position = position_dodge(0.1), size = .8) +
                ylim(1, 5) +
                ggtitle("Norm Violation") +
                theme_bw() + 
                labs(y="Perception of Norm Violation", x = "Anger") +
                theme(plot.title = element_text(hjust = 0.5), text = element_text(size = 10)) + 
                      scale_color_manual(values=met.brewer("Nizami", 2, direction=-1))      

    norms_p
      #ggsave(plot = norms_p, width = 5, height = 5, dpi = 300, filename = "norms_p.png")
```