---
title: "Analyses of the Perceptions of the Person/Entity who Caused Anger"
author: "Pooya Razavi"
date: "last knitted: `r Sys.time()`"
output: 
  html_document:
    theme: cosmo
    highlight: textmate
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---

This script represents the test of the hypotheses about the person or entity that caused the participant to experience anger, and how the different aspects of the behavior vary between justified and unjustified anger prototypes.

```{r setup, include=FALSE, warning=FALSE}
#load libraries
package_list <- c("dplyr", "tidyr", "ggplot2", "MetBrewer")
lapply(package_list, require, character.only = TRUE)

#read in the data
df <- readxl::read_xlsx("C:/Users/pooya/Dropbox (University of Oregon)/Anger Dissertation/Prototype study analysis/ProcessedData_F21_W22_S22_F22.xlsx")

#Function to report correlation
cor_report <- function(cor_output) {
          r <- cor_output[["estimate"]] %>% round(2)
          df <- cor_output[["parameter"]] %>% round(1)
          ci_lb <- (cor_output[["conf.int"]])[1] %>% round(2)
          ci_ub <- (cor_output[["conf.int"]])[2] %>% round(2)
          original_p <- cor_output[["p.value"]] %>% round(3)
          p <- if_else(original_p >= .001, paste0("= ", as.character(original_p)), "< .001")
          
          print(paste0("r(", df, ") = ", r, " [", ci_lb, ", ", ci_ub, "], p ", p))
    }

#Function to report independent-samples t-test
  ind_ttest_report <- function(iv, dv) {
      ttest <- t.test(dv ~ iv)
      effect_size <- effectsize::cohens_d(dv ~ iv, pooled_sd = FALSE)
      t <- ttest[["statistic"]] %>% round(2)
      df <- ttest[["parameter"]] %>% round(1)
      original_p <- ttest[["p.value"]] %>% round(3)
      p <- if_else(original_p >= .001, paste0("= ", as.character(original_p)), "< .001")
      d <- effect_size[1,1] %>% round(2)    
      
      print(paste0("t(", df, ") = ", t, ", p ", p, ", d = ", d))
      
  }
  
#Function to report paired-samples t-test
    paired_ttest_report <- function(t1, t2) {
      ttest <- t.test(Pair(t1, t2) ~ 1)
      effect_size <- effectsize::cohens_d(Pair(t1, t2) ~ 1, pooled_sd = FALSE)
      t <- ttest[["statistic"]] %>% round(2)
      df <- ttest[["parameter"]] %>% round(1)
      original_p <- ttest[["p.value"]] %>% round(3)
      p <- if_else(original_p >= .001, paste0("= ", as.character(original_p)), "< .001")
      d <- effect_size[1,1] %>% round(2)    
      
      print(paste0("t(", df, ") = ", t, ", p ", p, ", d = ", d))
      
  }
  
  
knitr::opts_chunk$set(echo = TRUE)
```


```{r, data-exclusion}
#assigning values to factor levels
  df$NarrativeWritten <- as.factor(df$NarrativeWritten)
  df$NarrativeRelevant <- as.factor(df$NarrativeRelevant)
  df$Condition <- as.factor(df$Condition)
  
  levels(df$NarrativeWritten) <- c("No", "Yes")
  levels(df$NarrativeRelevant) <- c("No", "Yes", NA, NA) 
  levels(df$Condition) <- c("justified", "nonjustified", NA)

#drop cases following preregistration
  df1 <- df %>% 
    filter(NarrativeWritten != "No") %>% 
    filter(NarrativeRelevant != "No") %>% 
    filter(!is.na(Condition))

```

# Causal Attribution

**Hypothesis:** In anger events that are perceived as justified (compared to the unjustified ones), participants are more likely to consider the cause of anger to stem from the target’s internal and stable characteristics (as opposed to the external and changeable circumstances).

```{r, causal-attr}
#First, test the correlation between the two items: 
cor.test(df1$cause_circumst, df1$behave_same) %>% cor_report()


#Since the items are not strongly correlated, a composite score will not be created. Conditions will be compared for each item separately:
  #cause: internal vs. external 
    df1 %>% 
        t.test(cause_circumst ~ Condition, data = .)
    df1 %>% 
        effectsize::cohens_d(cause_circumst ~ Condition, data = .)
    #For reporting:
      ind_ttest_report(df1$Condition, df1$cause_circumst)
    
  #cause: stable vs. the same 
    df1 %>% 
        t.test(behave_same ~ Condition, data = .)
    df1 %>% 
        effectsize::cohens_d(behave_same ~ Condition, data = .)
    #For reporting
      ind_ttest_report(df1$Condition, df1$behave_same)

```

### Moral Character

**Hypothesis:** In anger events perceived as justified (compared to the unjustified ones), the target is more likely to be seen as having weaker moral and ethical values. 

```{r, moral-char}
#create the morality composite score using the 10 items from Walker & Pitts (1998) 

  df_morality <- df1 %>% 
                    select(ResponseId, Condition, starts_with("wp_"), gw_honest, gw_principled)

  keys.list <- list(highly_moral=c("wp_concerned_right","wp_faithful","wp_clear_values", "wp_lawabiding", 
                                   "wp_strong_beliefs", "wp_distinguishes", "wp_dev_conscience", "wp_ethical", 
                                   "gw_honest", "gw_principled"))

  morality_scores <- psych::scoreItems(keys.list, df_morality)
  morality_scores
  df_morality <- cbind(df_morality, morality_scores$scores)

#Compare conditions
    df_morality %>% 
        t.test(highly_moral ~ Condition, data = .)
    df_morality %>% 
        effectsize::cohens_d(highly_moral ~ Condition, data = .)
      #For reporting
          ind_ttest_report(df_morality$Condition, df_morality$highly_moral)
```

# Moral-Relational Judgments

**Research Question:** What are the differences between justified vs. unjustified anger in terms of the moral-relational judgments of the target (as measured using Goodwin et al.'s [2014] trait characteristic items)? <br>

Goodwin et al. (2014) demonstrate that judgments of morality and warmth are separable and can provide unique informational value for person perception. To examine the perceptions of the target along these two critical dimensions, dimensionality reduction analysis (i.e., PCA) will be conducted on the character judgments from Goodwin et al. (2014). <br>

The number of components will be decided based on the scree plot, parallel analysis, and Velicor's MAP. If multiple alternative solutions are suggested by these methods, PCA will be conducted for all the alternatives number of components, and the optimal solution will be chosen based on component interpretability. <br>

## Component Structure

```{r, moral-relational-pca}
df_moral_relational <- df1 %>% 
                    select(ResponseId, Condition, starts_with("gw_"))

#To make a decision about the number of components
  #scree plot
  df_moral_relational %>% 
    select(starts_with("gw_")) %>% 
    psych::scree(hline = -1) #4 components
  
  #parallel analysis
  df_moral_relational %>% 
    select(starts_with("gw_")) %>% 
    psych::fa.parallel() #3 components
  
  #Velicor's MAP
  df_moral_relational %>% 
    select(starts_with("gw_")) %>% 
    psych::nfactors(n = 32) #4 components

#PCA with 3 components (based on the outcome of the parallel analyses)
  pca_3_component <- df_moral_relational %>% 
                        select(starts_with("gw_")) %>%  
                        psych::principal(., nfactors = 3, rotate = "varimax") 
        
  three_comp_outcome <- psych::kaiser(pca_3_component, rotate = "Varimax") %>% psych::fa.sort()

  three_comp_outcome[["loadings"]] %>% 
                    as.data.frame() %>% 
                    tibble::rownames_to_column("item") %>% 
                    mutate(Item = gsub("^.{0,3}", "", item)) %>% 
                    select(Item, starts_with("RC")) %>% 
                    knitr::kable(digits = 2) %>%
                    kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>% 
                    kableExtra::kable_paper(full_width = F)

  #This solution has a large proportion of cross-loading items, especially on the 2nd and 3rd components.
  
#PCA with 4 components (based on the outcome of the scree plot and the Velicor's MAP)
  pca_4_component <- df_moral_relational %>% 
                        select(starts_with("gw_")) %>%  
                        psych::principal(., nfactors = 4, rotate = "varimax") 
        
  four_comp_outcome <- psych::kaiser(pca_4_component, rotate = "Varimax") %>% psych::fa.sort()

  four_comp_outcome[["loadings"]] %>% 
                    as.data.frame() %>% 
                    tibble::rownames_to_column("item") %>% 
                    mutate(Item = gsub("^.{0,3}", "", item)) %>% 
                    select(Item, starts_with("RC")) %>% 
                    knitr::kable(digits = 2) %>%
                    kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>% 
                    kableExtra::kable_paper(full_width = F)  
  
  
  #find the cross-loading items
  four_comp_outcome[["loadings"]] %>% 
                    as.data.frame() %>% 
                    tibble::rownames_to_column("item") %>% 
                    mutate(Item = gsub("^.{0,3}", "", item)) %>% 
                    select(Item, starts_with("RC")) %>% 
                    mutate(RC1_RC2 = if_else(abs(RC1-RC2) > .1, " ", "*CL*"),
                           RC1_RC3 = if_else(abs(RC1-RC3) > .1, " ", "*CL*"),
                           RC1_RC4 = if_else(abs(RC1-RC4) > .1, " ", "*CL*"),
                           RC2_RC3 = if_else(abs(RC2-RC3) > .1, " ", "*CL*"),
                           RC2_RC4 = if_else(abs(RC2-RC4) > .1, " ", "*CL*"),
                           RC3_RC4 = if_else(abs(RC3-RC4) > .1, " ", "*CL*")) %>% 
                    knitr::kable(digits = 2) %>%
                    kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>% 
                    kableExtra::kable_paper(full_width = T) 
  
  #"Principled" has the strongest cross-loading. It will be dropped before the next iteration.
  pca_4_component_i2 <- df_moral_relational %>% 
                        select(starts_with("gw_"), -gw_principled) %>%  
                        psych::principal(., nfactors = 4, rotate = "varimax") 
        
  four_comp_outcome_i2 <- psych::kaiser(pca_4_component_i2, rotate = "Varimax") %>% psych::fa.sort()
  
  four_comp_outcome_i2[["loadings"]] %>% 
                    as.data.frame() %>% 
                    tibble::rownames_to_column("item") %>% 
                    mutate(Item = gsub("^.{0,3}", "", item)) %>% 
                    select(Item, starts_with("RC")) %>% 
                    mutate(RC1_RC2 = if_else(abs(RC1-RC2) > .1, " ", "*CL*"),
                           RC1_RC3 = if_else(abs(RC1-RC3) > .1, " ", "*CL*"),
                           RC1_RC4 = if_else(abs(RC1-RC4) > .1, " ", "*CL*"),
                           RC2_RC3 = if_else(abs(RC2-RC3) > .1, " ", "*CL*"),
                           RC2_RC4 = if_else(abs(RC2-RC4) > .1, " ", "*CL*"),
                           RC3_RC4 = if_else(abs(RC3-RC4) > .1, " ", "*CL*")) %>% 
                    knitr::kable(digits = 2) %>%
                    kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>% 
                    kableExtra::kable_paper(full_width = T) 
  
  #"Courageous" has the strongest cross-loading. It will be dropped before the next iteration.
  pca_4_component_i3 <- df_moral_relational %>% 
                        select(starts_with("gw_"), -gw_principled, -gw_courageous) %>%  
                        psych::principal(., nfactors = 4, rotate = "varimax") 
        
  four_comp_outcome_i3 <- psych::kaiser(pca_4_component_i3, rotate = "Varimax") %>% psych::fa.sort()
  
  four_comp_outcome_i3[["loadings"]] %>% 
                    as.data.frame() %>% 
                    tibble::rownames_to_column("item") %>% 
                    mutate(Item = gsub("^.{0,3}", "", item)) %>% 
                    select(Item, starts_with("RC")) %>% 
                    mutate(RC1_RC2 = if_else(abs(RC1-RC2) > .1, " ", "*CL*"),
                           RC1_RC3 = if_else(abs(RC1-RC3) > .1, " ", "*CL*"),
                           RC1_RC4 = if_else(abs(RC1-RC4) > .1, " ", "*CL*"),
                           RC2_RC3 = if_else(abs(RC2-RC3) > .1, " ", "*CL*"),
                           RC2_RC4 = if_else(abs(RC2-RC4) > .1, " ", "*CL*"),
                           RC3_RC4 = if_else(abs(RC3-RC4) > .1, " ", "*CL*")) %>% 
                    knitr::kable(digits = 2) %>%
                    kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>% 
                    kableExtra::kable_paper(full_width = T) 
  
  #"Warm" has the strongest cross-loading. It will be dropped before the next iteration.
  pca_4_component_i4 <- df_moral_relational %>% 
                        select(starts_with("gw_"), -gw_principled, -gw_courageous, -gw_warm) %>%  
                        psych::principal(., nfactors = 4, rotate = "varimax") 
        
  four_comp_outcome_i4 <- psych::kaiser(pca_4_component_i4, rotate = "Varimax") %>% psych::fa.sort()
  
  four_comp_outcome_i4[["loadings"]] %>% 
                    as.data.frame() %>% 
                    tibble::rownames_to_column("item") %>% 
                    mutate(Item = gsub("^.{0,3}", "", item)) %>% 
                    select(Item, starts_with("RC")) %>% 
                    mutate(RC1_RC2 = if_else(abs(RC1-RC2) > .1, " ", "*CL*"),
                           RC1_RC3 = if_else(abs(RC1-RC3) > .1, " ", "*CL*"),
                           RC1_RC4 = if_else(abs(RC1-RC4) > .1, " ", "*CL*"),
                           RC2_RC3 = if_else(abs(RC2-RC3) > .1, " ", "*CL*"),
                           RC2_RC4 = if_else(abs(RC2-RC4) > .1, " ", "*CL*"),
                           RC3_RC4 = if_else(abs(RC3-RC4) > .1, " ", "*CL*")) %>% 
                    knitr::kable(digits = 2) %>%
                    kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>% 
                    kableExtra::kable_paper(full_width = T) 
  #"Agreeable" is the only cross-loading item. It will be dropped before the next iteration.
  pca_4_component_i5 <- df_moral_relational %>% 
                        select(starts_with("gw_"), -gw_principled, -gw_courageous, -gw_warm, -gw_agreeable) %>%  
                        psych::principal(., nfactors = 4, rotate = "varimax") 
        
  four_comp_outcome_i5 <- psych::kaiser(pca_4_component_i5, rotate = "Varimax") %>% psych::fa.sort()
  
  four_comp_outcome_i5[["loadings"]] %>% 
                    as.data.frame() %>% 
                    tibble::rownames_to_column("item") %>% 
                    mutate(Item = gsub("^.{0,3}", "", item)) %>% 
                    select(Item, starts_with("RC")) %>% 
                    mutate(RC1_RC2 = if_else(abs(RC1-RC2) > .1, " ", "*CL*"),
                           RC1_RC3 = if_else(abs(RC1-RC3) > .1, " ", "*CL*"),
                           RC1_RC4 = if_else(abs(RC1-RC4) > .1, " ", "*CL*"),
                           RC2_RC3 = if_else(abs(RC2-RC3) > .1, " ", "*CL*"),
                           RC2_RC4 = if_else(abs(RC2-RC4) > .1, " ", "*CL*"),
                           RC3_RC4 = if_else(abs(RC3-RC4) > .1, " ", "*CL*")) %>% 
                    knitr::kable(digits = 2) %>%
                    kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>% 
                    kableExtra::kable_paper(full_width = T) 

  #to save the output in a .csv (for the manuscript table)
  #four_comp_outcome_i5[["loadings"]] %>% 
  #                  as.data.frame() %>% 
  #                  tibble::rownames_to_column("item") %>% 
  #                  mutate(Item = gsub("^.{0,3}", "", item)) %>% 
  #                  select(Item, starts_with("RC")) %>% write.csv("moral_relational_pca.csv")
```

Create scree plot, parallel analyses, and Velicor's MAP figures for the manuscript:

```{r}
#figure for the manuscript
  df_moral_relational %>% 
            select(starts_with("gw_")) %>% 
            psych::scree(factors = FALSE, hline = .73)
  df_moral_relational %>% 
                      select(starts_with("gw_")) %>% 
                      psych::fa.parallel(fa = "pc")
  
  velicor_map <- df_moral_relational %>% 
                      select(starts_with("gw_")) %>% 
                      psych::nfactors(n = 32)
  plot(velicor_map[["map"]],
       main = "Velicor's MAP",
       xlab = "Number of Factos",
       ylab = "MAP Index",
       xlim=c(0,25), ylim=c(0,0.10),
       type = "b")
  abline(h=0.013, col="blue")
```


Based on the PCA results, scale scores will be created. 

```{r, mora;-relational-scores}
  keys.list <- list(communal=c("gw_trustworthy","gw_empathetic","gw_kind","gw_forgiving","gw_fair","gw_humble","gw_loyal","gw_giving","gw_grateful","gw_just","gw_honest","gw_cooperative","gw_helpful"),
                    sociable=c("gw_sociable","gw_enthusiastic","gw_happy","gw_playful","gw_easygoing","gw_funny","gw_athletic"),
                    smart=c("gw_organized","gw_logical","gw_responsible","gw_intelligent","gw_clever"),
                    creative=c("gw_musical","gw_creative","gw_innovative"))

  moral_relation_scores <- psych::scoreItems(keys.list, df_moral_relational)
  moral_relation_scores
  df_moral_relational <- cbind(df_moral_relational, moral_relation_scores$scores)
  
  #The 3 lines below will save the APA-style correlation table in a word document:
  #df_moral_relational %>% 
  #  select(communal, sociable, smart, creative) %>% 
  #  apaTables::apa.cor.table(.,filename = "moral_relational_cor.doc", table.number = 1, show.conf.interval = T)
  
```

## Comparison of Anger Variants

```{r, moral-relational-compare}
#descriptives  
  df_moral_relational %>% 
    group_by(Condition) %>% 
    select(communal, sociable, smart, creative) %>% 
    summarise_all(list(mean, sd))

#Communal
    df_moral_relational %>% 
        t.test(communal ~ Condition, data = .)
    df_moral_relational %>% 
        effectsize::cohens_d(communal ~ Condition, data = .)
    #For reporting
    ind_ttest_report(df_moral_relational$Condition, df_moral_relational$communal)

#Sociable
    df_moral_relational %>% 
        t.test(sociable ~ Condition, data = .)
    df_moral_relational %>% 
        effectsize::cohens_d(sociable ~ Condition, data = .)
    #For reporting
    ind_ttest_report(df_moral_relational$Condition, df_moral_relational$sociable)

#Smart
    df_moral_relational %>% 
        t.test(smart ~ Condition, data = .)
    df_moral_relational %>% 
        effectsize::cohens_d(smart ~ Condition, data = .)
    #For reporting
    ind_ttest_report(df_moral_relational$Condition, df_moral_relational$smart)
    
#Creative
    df_moral_relational %>% 
        t.test(creative ~ Condition, data = .)
    df_moral_relational %>% 
        effectsize::cohens_d(creative ~ Condition, data = .)
    #For reporting
    ind_ttest_report(df_moral_relational$Condition, df_moral_relational$creative)    


```

